On Developing User Interfaces for Piloting Unmanned Systems

### 2014-02-19 Papers/RSN2014/paper.tex
* Updated text of workshop paper

%---------------------------------------------------------------------%
\section{sec:introduction}

@@

In a waypoint navigation system, the user must select a device in an interface
and then select a waypoint to traverse to. Most of these interfaces allow users
to queue up waypoints (essentially plot a path of points), which allows an
operator to plan out complicated missions. Figure ~\ref{fig:guiding} lays out
the mission we asked of three volunteers, whom we allowed to replot their
missions after seeing results that displayed the required clicks and the area
that actually ended up being covered. The larger area is the main search area
and the smaller areas are priority regions, places that are more important due
to the mission and must be searched quickly (e.g., in a search-and-rescue
mission where known survivors are at and triage decisions need to be made).

@@

The following figures (Figure ~\ref{fig:user1_1}, 
~\ref{fig:user2_2}, ~\ref{fig:user3_2})
are the results of handing the waypoint navigation to random
people in our research lab, one of whom had experience flying waypoint systems.
For space reasons, we picked only the three that showed the most interesting
characters--whether showing higher user error or excessive diligence.

These are by no means meant to categorically and exhaustively explore the space
of user interactions possible in the system. We show these here to highlight the
differences between the mission requirements and the kinds of errors that are likely with
human usage. Additionally, after these images, we breakdown the average user
actions required to send one drone on a mission like this. As swarms grow larger,
the waypoint system becomes even more cumbersome to use.

@@

Table ~\ref{tab:averagewaypoints} shows the marker counts, readjustments,
user actions and time taken to use the waypoint navigation interface to search these
three areas. The time taken and user actions varied wildly, and the accuracy of the
search was also highly variable. These results highlight some of the problems in
using such waypoint-based systems in mission-critical contexts like
search-and-rescue but also in agriculture or any number of other scenarios where
accuracy is important. A
UAV attempting to spray pesticides over a field and missing large patches can
be disasterous for crop growth and maintenance. A UAV that does not locate a
survivor could result in unnecessary fatalities and improper triage in a disaster
area.

@@

In the next section, we will discuss the results of using the GAMS
interface on similar problems.

%---------------------------------------------------------------------%
\section{The GAMS Interface and Software}
\label{sec:gams_interface}

@@

Primitive operations are functions like takeoff, land, move-to-gps, and other
operations that are considered primitive AI. Network bridging is an instruction
to the drones to move to locations that allow network traffic to pass between
locations. Area coverage is a distributed AI that causes the drones to collaborate
to cover a single rectangular area. Prioritized area coverage is a distributed AI
that allows for the designation of prioritized regions within a main area coverage
that the UAS will try to cover first.

The underlying artificial intelligence is implemented in the MADARA Knowledge and
Reasoning Engine (KaRL) engine ~\cite{Edmondson-KaRL-ksem2011}, a specialized
and fast, real-time reasoning engine that was built for contextual inferencing in
distributed, decentralized sensor networks. Our interface, which allows
interaction with this AI, is built in Java for any device running the Android
operating system. This allows us to control a swarm from a tablet or smartphone.
Both the interfaces and the AI is released as open source under a BSD license
at http://gams-cmu.googlecode.com.

In the next section, we will detail a small series of experiments that showcase the
power of the interface and how it improves on the state-of-the-art in both required
user actions and time taken to input a new swarm mission.

@@

The major takeaways from this data are that the waypoint system interface, even
with just two drones, requires more user actions than the GAMS interface.
Additionally, as shown in Section~\ref{sec:waypoints}, waypoint systems are prone to
error, and the more error inserted into guessing the sensor radius, the more actions
required and the more the area covered has overlaps or gaps. In the expressed data,
we do not require additional user actions ($\gamma$) to fix gaps in the area covered
by waypoint navigation systems.

@@

In this paper, we have discussed insights into research conducted in the Group
Autonomy for Mobile Systems project at Carnegie Mellon University in regards
to user interfaces for swarms of unmanned autonomous systems. The user
interfaces developed and showcased here have been shown to decrease
user actions required and also reduce error by relying on autonomy within
the devices to cover areas correctly. Future work on this project is directed toward
using these interfaces to direct dozens of UAS to perform complex mission tasks
for search-and-rescue, agriculture, border patrol, and other tasks.

>>>>

%---------------------------------------------------------------------%
\section{Introduction}
\label{sec:introduction}

@@

Imagine a search-and-rescue situation during an earthquake where two large
buildings have collapsed and rescuers need to search the area for survivors. We
highlight such a situation in Figure~\ref{fig:guiding}. Survivors could be located
throughout the affected area (the larger region) but are more likely to be found
in the collapsed building locations, specified as the smaller rectangles. As part
of a simple experiment, we offered a waypoint interface with this overlay, the
sensor range of a small UAV, and asked them to search the entire area with
a preference on the building locations. Each user was allowed to redo their
waypoints after an initial attempt, and we kept track of the number of actions
required, the unexplored areas, and the time required to input the mission.

@@

Figures~\ref{fig:user1_1}, ~\ref{fig:user1_2}
~\ref{fig:user2_2}, and ~\ref{fig:user3_2}) show the results of these
simple experiments with three random persons in our lab, only one of whom
had any experience with waypoint systems. Even when given extra attempts
and allowed time to observe the map and mission with ample time before
moving UAVs around the map, human operators required dozens of actions,
inserted human error, and took dozens of seconds to tell even a single UAV
how to search the area.

@@

Table ~\ref{tab:averagewaypoints} shows the marker counts, readjustments,
user actions and time taken to use the waypoint navigation interface to search these
three areas. The time taken and user actions varied wildly, and the accuracy of the
search was also highly variable. These results highlight some of the problems in
using such waypoint-based systems in mission-critical contexts like
search-and-rescue but are also applicable in agriculture, where spraying the wrong
field can result in the loss of a crop, or any number of other scenarios where
accuracy and timing are important.

@@

In the next section, we will discuss the development of user interfaces and
coupled artificial intelligence to simplify interactions with one or more robotic
systems.

%---------------------------------------------------------------------%
\section{The GAMS Interface and Software}
\label{sec:gamsinterface}

@@

Primitive operations are functions like takeoff, land, move-to-gps, and other
operations that are considered primitive AI. Network bridging is an instruction
to the UAS to move to locations that allow network traffic to pass between
locations. Area coverage is a distributed AI that instructs the UAS to collaborate
in order to cover a single rectangular area. Prioritized area coverage is a distributed AI
that provides for prioritized regions within a main area coverage
that the UAS will try to cover first.

The underlying artificial intelligence is implemented in the MADARA Knowledge and
Reasoning Engine (KaRL) engine ~\cite{Edmondson-KaRL-ksem2011}, a specialized
and fast, real-time reasoning engine that was built for contextual inferencing in
distributed, decentralized sensor networks. Our control interface is built for Google's
Android Operating System~\cite{nimodia2012android}, which allows us to control a swarm from any Google
tablet or smartphone. Both the interfaces and the AI is released as open source under a BSD license
at http://gams-cmu.googlecode.com.

In the next section, we will detail a small series of experiments that showcase the
power of the interface and how it improves on the state-of-the-art in reducing the
required user actions and consequently the time taken to input a new swarm mission.

@@

The major takeaways from this data are that the waypoint system interface, even
with just two drones, requires more user actions than the GAMS interface.
Additionally, as shown in Section~\ref{sec:waypoints}, waypoint systems are prone to
error, and the more error inserted into guessing the sensor radius, the more actions
required and the more the area covered has overlaps or gaps. In the expressed data,
we do not require additional user actions ($\gamma$) to fix gaps in the area covered
by waypoint navigation systems. In the real world, unexplored areas would have to
be retasked, which will result in more time taken and focus by the human operator.
Retasking for missed areas would also increase the number of required user actions.

@@

In this paper, we have discussed insights into research conducted in the Group
Autonomy for Mobile Systems project at Carnegie Mellon University in regards
to user interfaces for swarms of unmanned autonomous systems. The user
interfaces described here have been shown to decrease
user actions required and also reduce error by relying on autonomy within
the devices to cover areas correctly. Future work on this project is directed toward
using these interfaces to direct dozens of UAS to perform complex mission tasks
for search-and-rescue, agriculture, border patrol, and other tasks.
